---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# addreg

[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/addreg)](https://cran.r-project.org/package=addreg)

`addreg` provides methods for fitting identity-link GLMs and GAMs, using EM-type algorithms with more stable convergence properties than standard methods.

An example of periodic non-convergence using `glm` (run with `trace = TRUE` to see deviance at each iteration):

```{r, echo = TRUE, warning = FALSE}
require(glm2, quietly = TRUE)
data(crabs)

crabs.boot <- crabs[crabs$Rep1, -c(5:6)]

t.glm <- system.time(
  fit.glm <- glm(Satellites ~ Width + Dark + GoodSpine, data = crabs.boot, family = poisson(identity),
                    start = rep(1, 4), maxit = 500)
)
```

The combinatorial EM method (Marschner, 2010) provides stable convergence:
```{r, echo = TRUE, message = FALSE}
require(addreg, quietly = TRUE)
t.cem <- system.time(
  fit.cem <- addreg(Satellites ~ Width + Dark + GoodSpine, data = crabs.boot, family = poisson,
                    start = rep(1, 4))
)
```

...but it can take a while. Using an overparameterised EM approach removes the need to run $2^3 = 8$ separate EM algorithms:
```{r echo = TRUE, message = FALSE}
t.em <- system.time(fit.em <- update(fit.cem, method = "em"))
```

while generic EM acceleration algorithms from the `turboEM` package --- implemented in version $\geq$ 3.0 --- can speed this up further still:
```{r, echo = TRUE, message = FALSE}
t.cem.acc <- system.time(fit.cem.acc <- update(fit.cem, accelerate = "squarem"))
t.em.acc <- system.time(fit.em.acc <- update(fit.em, accelerate = "squarem"))
```

Comparison of results:
```{r, echo = FALSE}
fit.list <- list(fit.glm, fit.cem, fit.em, fit.cem.acc, fit.em.acc)
time.list <- list(t.glm, t.cem, t.em, t.cem.acc, t.em.acc)
res <- data.frame(converged = sapply(fit.list, function(x) x$converged),
                  logLik = sapply(fit.list, logLik),
                  iterations = sapply(fit.list, function(x) x$iter[1]),
                  time = sapply(time.list, function(x) x[3]))
rownames(res) <- c("glm", "cem", "em", "cem.acc", "em.acc")
res
```

The combinatorial EM algorithms for identity-link binomial (Donoghoe and Marschner, 2014), negative binomial (Donoghoe and Marschner, 2016) and gamma models are also available, using `family = binomial`, `family = negbin1` and `family = gamma1`, respectively.

Semi-parametric regression using B-splines (Donoghoe and Marschner, 2015) can be incorporated by using the `addreg.smooth` function. See `example(addreg.smooth)` for a simple example.

## Installation

Get the released version from CRAN:

```{r, echo = TRUE, eval = FALSE}
install.packages("addreg")
```

Or the development version from github:

```{r, echo = TRUE, eval = FALSE}
# install.packages("devtools")
devtools::install_github("mdonoghoe/addreg")
```

## References
* Donoghoe, M. W. and I. C. Marschner (2014). Stable computational methods for additive binomial models with application to adjusted risk differences. _Computational Statistics and Data Analysis_ __80__: 184-196.
* Donoghoe, M. W. and I. C. Marschner (2015). Flexible regression models for rate differences, risk differences and relative risks. _International Journal of Biostatistics_ __11__(1): 91-108.
* Donoghoe, M. W. and I. C. Marschner (2016). Estimation of adjusted rate differences using additive negative binomial regression. _Statistics in Medicine_ __35__(18): 3166-3178.
* Marschner, I. C. (2010). Stable computation of maximum likelihood estimates in identity link Poisson regression. _Journal of Computational and Graphical Statistics_ __19__(3): 666-683.